# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv", header=TRUE)
activity_clean <- activity[which(activity$steps != "NA"),]
```
## What is mean total number of steps taken per day?
```{r}
library(plyr)
stepsperday <- ddply(activity_clean, .(date), summarise, steps=sum(steps))
hist(stepsperday$steps, main="Number of Steps",xlab="Total number of steps taken each day", col="red")

mean(stepsperday$steps)
median(stepsperday$steps)
```
## What is the average daily activity pattern?
```{r}
avgdailypattern <- tapply(activity$steps, activity$interval, mean, na.rm = T)
plot(avgdailypattern, type = "l", main = ("Steps vs. Interval (daily average)"), ylab = "# of steps")

seq(along = avgdailypattern)[avgdailypattern == max(avgdailypattern)]
```
## Imputing missing values
The mean of the the intervals was used as a proxy for the missing data.

```{r}
sum(as.numeric(is.na(activity$steps)))

avgdailypatternvector <- as.vector(avgdailypattern)
avgdailypatternvector <- rep(avgdailypatternvector, 61)
avgdailypatternvector[!is.na(activity$steps)] = 1
missingdatavector <- as.vector(activity$steps)
missingdatavector[is.na(missingdatavector)] = 1
avgforna_activity <- activity
avgforna_activity$steps <- avgdailypatternvector * missingdatavector

complete_stepsperday <- ddply(avgforna_activity, .(date), summarise, steps=sum(steps))

hist(complete_stepsperday$steps, main="Number of Steps", xlab="Total number of steps taken each day", col="red",)
```
The added mean values have increased the amount of steps in the middle of the histogram.

```{r}
mean(complete_stepsperday$steps)
median(complete_stepsperday$steps)

```
And the median now is equal to the mean due to the added mean values.

## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lattice)
weekdays <- weekdays(as.Date(activity_clean$date))
data_with_weekdays <- transform(activity_clean, day=weekdays)
data_with_weekdays$wk <- ifelse(data_with_weekdays$day %in% c("Saturday", "Sunday"),"weekend", "weekday")
average_by_interval_wk <- ddply(data_with_weekdays, .(interval, wk), summarise, steps=mean(steps))

xyplot(steps ~ interval | wk, data = average_by_interval_wk, layout = c(1, 2), type="l", xlab = "Interval", ylab = "Number of steps")

```
